\section{The \algname{} Algorithm}\label{sec:algorithm}

We describe \algname{} (\emph{Adaptive Symbolic Importance
  Partitioning}).  \Cref{sec:frontier} introduces the frontier data
structure.  \Cref{sec:hw-decomp} states the central decomposition
theorem, which is the load-bearing correctness result of the paper.
\Cref{sec:asip-loop} gives the driver loop.

\subsection{The Frontier}\label{sec:frontier}

\algname{} maintains a \emph{frontier}: a tree of path-condition
regions $\{R_\pi\}_{\pi \in \partset}$ whose leaves form a partition
of $\inX$.  Each leaf $\pi$ carries four pieces of data:
\begin{itemize}
\item a symbolic region formula $F_\pi$ in the theory of
  linear integer/bitvector arithmetic extended with equality and
  propositional connectives;
\item a probability estimate $\hwhat_\pi \approx \Pr_\dist[X \in R_\pi]$;
\item a sample-based hit-rate estimate
  $\hpathat_\pi \approx \Pr[\prop(\prog(X))=1 \mid X \in R_\pi]$;
\item a lifecycle status in
  $\{\mathsf{OPEN}, \mathsf{CLOSED\text{-}T},
    \mathsf{CLOSED\text{-}F}, \mathsf{EMPTY}\}$.
\end{itemize}

A leaf is \textsc{open} if its hit rate is not yet symbolically
certified.  It is \textsc{closed-t} or \textsc{closed-f} if the SMT
backend has proved that $\prop(\prog(x)) = 1$ or $\prop(\prog(x)) = 0$
holds uniformly on the region.  It is \textsc{empty} if SMT has shown
$F_\pi \equiv \bot$.  The overall reliability estimator is
\begin{equation}\label{eq:muhat-def}
  \muhat \;=\; \sum_{\pi \in \partset} \hwhat_\pi \cdot \hpathat_\pi.
\end{equation}

\begin{lemma}[Mass-conservative refinement]\label{lem:mass-conservation}
  When \algname{} refines a leaf $\pi$ on a clause $b$, producing
  children $\pi{\wedge}b$ and $\pi{\wedge}\neg b$, the children's
  estimated masses satisfy
  $\hwhat_{\pi \wedge b} + \hwhat_{\pi \wedge \neg b} = \hwhat_\pi$
  exactly (modulo importance-sampling noise on non-axis-aligned
  regions, which is bounded in \Cref{sec:implementation}).
\end{lemma}

\begin{proof}[Proof sketch]
  For axis-aligned LIA children, the closed-form mass of each child
  is a product of marginal-PMF differences over disjoint coordinate
  ranges, which sum exactly to the parent's mass.  For general
  regions, \algname{} draws one importance-sampling batch from
  $\dist \mid R_\pi$, partitions it by $b$, and splits $\hwhat_\pi$
  proportionally: $\hwhat_{\pi \wedge b} = \hwhat_\pi \cdot
  \hpathat_{b \mid \pi}$ and symmetrically for $\neg b$.  The split
  conserves $\hwhat_\pi$ by construction.
\end{proof}

\begin{lemma}[Closure soundness]\label{lem:closure-soundness}
  Suppose the SMT oracle returns \unsat{} for
  $F_\pi \wedge \neg\,\mathrm{path}_\pi$, where
  $\mathrm{path}_\pi$ is the conjunction of branch clauses observed
  at $\pi$.  Then every $x \in R_\pi$ traverses
  $\mathrm{path}_\pi$ in $\prog$.  If additionally every
  observation at $\pi$ agrees on the same $\prop$-value
  $v \in \{0,1\}$, then $\prop(\prog(x)) = v$ for every
  $x \in R_\pi$.
\end{lemma}

\begin{proof}[Proof sketch]
  Suppose some $x \in R_\pi$ takes a different path.  Then some
  branch clause $b_j$ of $\mathrm{path}_\pi$ is violated by $x$,
  so $x \models F_\pi \wedge \neg\,\mathrm{path}_\pi$, contradicting
  the oracle's \unsat{} verdict (by soundness of the SMT backend).
  Determinism of $\prog$ then forces $\prop(\prog(x)) = v$ uniformly
  on $R_\pi$.
\end{proof}

\subsection{The Central Decomposition Theorem}\label{sec:hw-decomp}

Let $\Pi_T$ denote the frontier at the algorithm's stopping time $T$,
decomposed as
$\Pi_T = \openleaves \cup \closedleaves \cup \Pi^{\mathsf{EMPTY}}_T$.
Define the two half-width terms:
\begin{align}
  \Wopen(T) \;&=\; \sum_{\pi \in \openleaves} \hwhat_\pi,\\
  \epsstat(T) \;&=\; \sum_{\pi \in \openleaves}
    \hwhat_\pi \cdot W_\pi(T),
\end{align}
where $W_\pi(T)$ is an anytime-valid $(1-\delta_\pi)$-half-width for
the per-leaf hit rate, with Bonferroni allocation
$\delta_\pi = \delta / K_T$ over the $K_T \le K_{\max}$ leaves
created by the algorithm.

\begin{theorem}[Central decomposition; coverage at $1-\delta$]%
  \label{thm:hw-decomp}
  Under Assumption~1, Lemma~\ref{lem:mass-conservation},
  Lemma~\ref{lem:closure-soundness}, soundness of the SMT oracle, and
  an anytime-valid per-leaf bound $W_\pi$,
  \begin{equation}
    \mathbb{P}\!\Big[\,\big|\muhat_T - \mutrue\big|
      \;\le\; \epsstat(T) + \Wopen(T)\,\Big]
    \;\ge\; 1 - \delta.
  \end{equation}
\end{theorem}

\begin{proof}[Proof sketch]
  Decompose the estimation error:
  \[
    \muhat_T - \mutrue
    = \sum_{\pi \in \openleaves}
        \!\bigl(\hwhat_\pi\hpathat_\pi - w_\pi p_\pi\bigr)
      + \sum_{\pi \in \closedleaves}
        \!\bigl(\hwhat_\pi - w_\pi\bigr)p_\pi.
  \]
  The closed-leaf term vanishes for axis-aligned closed leaves
  (by the closed-form mass) and is bounded by
  Lemma~\ref{lem:mass-conservation} otherwise.  Each open-leaf term
  splits into a deterministic part---bounded in absolute value by
  $\hwhat_\pi$, summing to $\Wopen(T)$---and a stochastic part
  bounded with probability $\ge 1-\delta_\pi$ by
  $\hwhat_\pi W_\pi(T)$.  A union bound over the $K_T$ leaves
  controls the total stochastic deviation by $\epsstat(T)$ at
  confidence $1-\delta$.  The anytime property of $W_\pi$ makes
  the union bound valid at the data-dependent stopping time $T$.
\end{proof}

Theorem~\ref{thm:hw-decomp} is the load-bearing result.  Every
implementation choice in \Cref{sec:implementation} is justified by
reference to it, and every action the algorithm takes reduces either
$\epsstat$ or $\Wopen$.  The reported interval is the two-sided
projection
\begin{equation}\label{eq:interval}
  [L, U] \;=\;
  \bigl[\max(0,\,\muhat_T - \epsstat(T) - \Wopen(T)),\;
        \min(1,\,\muhat_T + \epsstat(T) + \Wopen(T))\bigr].
\end{equation}

\subsection{The Driver Loop}\label{sec:asip-loop}

At each iteration, \algname{} considers two action families.
\begin{description}
\item[\textsc{Sample}$(\pi, b)$.]
  Draw $b$ additional iid samples from $\dist \mid R_\pi$, increment
  $n_\pi$ by $b$, update $\hpathat_\pi$ and $W_\pi$.
  Expected gain: $\Delta\epsstat \approx \hwhat_\pi\cdot
  (W_\pi(n_\pi) - W_\pi(n_\pi+b))$.
  Cost: $b$ program evaluations.

\item[\textsc{Refine}$(\pi, b)$.]
  Select a branch clause $b$ observed at $\pi$.  Split $\pi$ into
  $\pi{\wedge}b$ and $\pi{\wedge}\neg b$, distributing $\hwhat_\pi$
  per Lemma~\ref{lem:mass-conservation}, then immediately attempt
  closure on each child (Lemma~\ref{lem:closure-soundness}).
  Expected gain: $\Delta\Wopen \approx \hwhat_\pi \cdot
  \Pr[\text{at least one child closes}]$.
  Cost: one SMT call (split) plus two (closure checks).
\end{description}

The scheduler selects $\arg\max_a\,\mathrm{gain}(a)/\mathrm{cost}(a)$,
where cost is normalised to one concolic run per unit.  The algorithm
terminates when $\epsstat(T) + \Wopen(T) \le \varepsilon$, or when
one of three optional caps fires: sample budget exhausted, wall-clock
limit reached, or no action exceeds a minimum gain-per-cost floor.

\begin{figure}[t]
  \caption{\algname{} (Adaptive Symbolic Importance Partitioning)}
  \label{alg:asip}
  \hrule\vspace{0.4em}
  \textbf{Input:} program $\prog$, distribution $\dist$, property
    $\prop$, target $(\varepsilon, \delta)$, optional budget $B$.\\
  \textbf{Output:} estimator $\muhat$ and certified interval $[L,U]$.
  \begin{flushleft}
  \begin{tabular}{rl}
    1: & $\partset \gets \{\pi_0\}$ with $F_{\pi_0} = \top$,
         $\hwhat_{\pi_0} = 1$ \hspace{1.5em}// single \textsc{open}
         leaf at root\\
    2: & bootstrap $n_{\mathrm{boot}}$ samples from $\dist$ at
         $\pi_0$;\; call \textsc{TryClose}($\pi_0$)\\
    3: & \textbf{while} not \textsc{Terminated}() \textbf{do}\\
    4: & \quad $\mathcal{A} \gets
         \{\textsc{Sample}(\pi,b),\,\textsc{Refine}(\pi,b)
           : \pi \in \openleaves\}$\\
    5: & \quad $a^\star \gets
         \arg\max_{a \in \mathcal{A}}\,\mathrm{gain}(a)/\mathrm{cost}(a)$\\
    6: & \quad execute $a^\star$;\; update $\partset$\\
    7: & \quad call \textsc{TryClose}($\pi$) for every newly-affected
         leaf\\
    8: & \textbf{end while}\\
    9: & \textbf{return} $(\muhat,\,[L,U])$ per
         \eqref{eq:muhat-def} and~\eqref{eq:interval}\\
  \end{tabular}
  \end{flushleft}
  \vspace{0.2em}\hrule
\end{figure}

\algname{} is parametrised by three engineering choices: the
candidate-clause heuristic for \textsc{Refine}, the per-leaf bound
construction $W_\pi$, and the SMT backend.  Each choice is discussed
in \Cref{sec:implementation}.
