\section{Implementation}\label{sec:implementation}

\toolname{} is an open-source Python prototype
of approximately $5{,}000$ lines of code with $290+$ tests.  This
section covers the three engineering choices that turn
\Cref{alg:asip} into a working artifact: the concolic front-end, the
per-leaf bound $W_\pi$, and the SMT abstraction layer.

\subsection{Concolic Execution and Branch Capture}\label{sec:concolic}

A concolic run~\cite{sen2005cute,godefroid2005dart} executes $\prog(x)$
on a concrete input $x$ while simultaneously tracking the symbolic
conditions on $x$ that drove each branch taken.  \toolname{}
implements this via a tagged-integer wrapper: arithmetic operations on
tagged values propagate their symbolic representations, and comparisons
emit a branch record $(\text{clause},\text{direction})$ before
delegating to the concrete operator.

Each run produces the path condition (the conjunction of branch clauses
along the path taken), the property value $\prop(\prog(x))$, and the
concrete input.  Property evaluation is intentionally \emph{outside}
the tracer: $\prop$ is evaluated on the concrete output, and the
resulting Boolean is appended to the run record.  This division of
labour keeps path conditions small, since $\prop$ is typically a thin
projection (\code{$y \le k$}, \code{popcount($y$) $\ge 0$}) that
should not inflate the symbolic representation.

\subsection{The Per-Leaf Bound $W_\pi$}\label{sec:per-leaf-bound}

Theorem~\ref{thm:hw-decomp} requires an anytime-valid
$(1-\delta_\pi)$-half-width for each open leaf's hit rate.  \toolname{}
exposes four choices via the \code{method} parameter.

\paragraph{Wilson (\code{method="wilson"}).}
The classical fixed-$n$ Wilson score interval~\cite{wilson1927}.  Tight
on Bernoulli leaves at a pre-committed sample count; \emph{not} sound
under data-dependent stopping.  Appropriate only when the budget is
fixed in advance.

\paragraph{Anytime Wilson (\code{method="anytime"}).}
The Wilson interval evaluated at confidence $1 - \delta_n$ with
$\delta_n = 6\delta_\pi/(\pi^2 n^2)$, union-bounded over all $n$ via
$\sum_{n=1}^\infty 6/(\pi^2 n^2) = 1$.  Sound under data-dependent
stopping; loose by a $\sqrt{\log n}$ factor and a $\pi^2/6$
inflation.

\paragraph{Predictable-plug-in empirical Bernstein (\code{method="betting"}).}
Theorem~2 of Waudby-Smith and Ramdas~\cite{waudbysmith2024betting}.
For a Bernoulli history $X_1,\ldots,X_n$ at leaf $\pi$, the half-width
is
\begin{equation}\label{eq:prpl-eb}
  W_\pi \;=\;
    \frac{\log(2/\delta_\pi) + \sum_{i=1}^n v_i\,\psi_E(\lambda_i)}
         {\sum_{i=1}^n \lambda_i},
  \qquad
  \psi_E(\lambda) = \tfrac{-\log(1-\lambda) - \lambda}{4},
\end{equation}
with $v_i = 4(X_i - \hat\mu_{i-1})^2$ and predictable bet
\begin{equation}
  \lambda_t \;=\; \min\!\Bigg(
    \sqrt{\frac{2\log(2/\delta_\pi)}
              {\hat\sigma_{t-1}^2 \cdot t\log(t+1)}},\;c\Bigg),
  \quad c \in (0,1),
\end{equation}
where $\hat\mu_t$ and $\hat\sigma_t^2$ are Wilson-regularised running
mean and variance.  The default truncation is $c = 1/2$, as
recommended in~\cite[\S3.2]{waudbysmith2024betting}.  PrPl-EB is
anytime-valid, closed-form, and variance-adaptive: its asymptotic
width is $\sqrt{2\sigma^2\log(2/\delta_\pi)}$, matching Bennett's
inequality without requiring known variance.  This is the recommended
default for certified results and is strictly tighter than
\code{anytime} on low-variance leaves.

\paragraph{Classical and Maurer--Pontil Bernstein.}
Provided as \code{bernstein} and \code{empirical-bernstein} for
baseline comparison only; both are fixed-$n$ bounds and not sound
under data-dependent stopping.

\subsection{The Closure Rule}\label{sec:closure}

A leaf is closed only when the SMT backend certifies
path-determinism:
\begin{verbatim}
def try_close(node):
    if n_samples < min_samples: return False
    if not all observed paths identical: return False
    if not all observed phi values identical: return False
    if is_satisfiable(F_pi AND NOT observed_path) != "unsat":
        return False
    node.status = CLOSED_T if phi == 1 else CLOSED_F
    return True
\end{verbatim}
There is no sample-based fallback.  A leaf whose closure the SMT
backend cannot certify (\sat{} or \unknown{}) remains open and
contributes to $\Wopen$.  This makes the soundness contract a single
lemma (Lemma~\ref{lem:closure-soundness}), at the cost of wider
intervals on programs whose path structure is symbolically
intractable.

\subsection{SMT Abstraction}\label{sec:smt}

The closure check, region-emptiness check, and refinement-clause
selection all go through an \code{SMTBackend} protocol with three
implementations: a Z3 backend~\cite{demoura2008z3} for production use,
a \code{MockBackend} restricted to axis-aligned LIA that permits the
test suite to run without external dependencies, and a
\code{CachedBackend} wrapper that memoises solver queries (typical
cache hit rate: $40$--$60\%$ on the benchmark suite).  All backends
guarantee soundness of \sat/\unsat{}; returning \unknown{} is always
admissible.

\subsection{Refinement-Clause Selection}\label{sec:clause-selection}

\toolname{}'s \textsc{Refine} action requires selecting a branch clause
$b$ from the finite set harvested from observed paths at $\pi$.
\toolname{} scores each candidate by expected mass balance---intuitively,
the clause that partitions $\hwhat_\pi$ closest to 50/50, maximising
the probability that at least one child closes.  The heuristic is
deliberately conservative; a principled alternative based on Craig
interpolation is identified as the primary next development direction
(\Cref{sec:conclusion}).

\subsection{Reproducibility}\label{sec:repro}

\toolname{}'s source code, benchmark suite, and reproduction scripts
are available at \code{https://github.com/<anon>/dise} (URL anonymised
for review).  The Docker image and \code{scripts/reproduce.sh}
regenerate all results in \Cref{sec:results} in approximately
$30$--$60$ minutes on commodity hardware.  Seed control is
centralised: each \code{(method, benchmark, seed)} triple is fully
deterministic.
