\section{Conclusion}\label{sec:conclusion}

We have presented \toolname{}, an algorithm for distribution-aware
reliability estimation that maintains a frontier of path-condition
regions and at each step selects between sampling and SMT-driven
refinement under a unified gain-per-cost criterion.  The certified
half-width decomposes into a statistical component $\epsstat$, driven
down by sampling, and a structural component $\Wopen$, driven down
by symbolic refinement.  This decomposition---Theorem~\ref{thm:hw-decomp}---is
simultaneously the algorithm's correctness argument and its scheduler.
Two implementation choices distinguish \toolname{} from prior work:
the per-leaf statistical bound instantiates the closed-form
predictable-plug-in empirical-Bernstein confidence sequence of
Waudby-Smith and Ramdas, and the closure rule is single-branch
(SMT-only), yielding a one-lemma soundness contract with no
sample-based fallback.

\paragraph{Where \toolname{} is the right tool.}
The headline experiment of \Cref{sec:rare-event-scaling} maps out
\toolname{}'s sweet spot.  When (i) the input distribution
concentrates on path-deterministic regions, (ii) the bootstrap
observes at least one example from each reachable region, and (iii)
the SMT cost on the program's path conditions is moderate, \toolname{}
certifies $\mutrue$ to half-width $\varepsilon$ in a number of samples
that is $\Theta(1)$ in $\mutrue$---a regime where plain Monte Carlo
requires $\Theta(1/(\mutrue\varepsilon^2))$ samples, an asymptotic
advantage of orders of magnitude on rare events.

\paragraph{Limitations.}
We identify four.

\begin{enumerate}
\item \emph{Distribution class.}  The (D1) restriction excludes
  Bayes-net-structured and general discrete distributions.  Lifting it
  requires ancestor-conditioned constrained samplers and a mass
  estimator that handles coordinate dependence.
\item \emph{SMT cost on non-LIA path conditions.}
  \Cref{sec:suite} reports that on \code{gcd\_steps\_le\_5\_BG},
  \code{miller\_rabin\_w=2\_BG}, and similar non-LIA or BG-distributed
  benchmarks, the closure check saturates the wall-clock budget
  before $\Wopen$ shrinks meaningfully.
\item \emph{Refinement-clause selection.}  The current heuristic
  picks from clauses observed in samples; it does not synthesise new
  predicates.  Craig interpolation is the natural principled
  alternative.
\item \emph{Bootstrap blindness on very rare events.}
  When $\mutrue \le 10^{-4}$, the bootstrap is unlikely to observe a
  hit in the rare region; the SMT closure check on the dominant
  \textsc{closed-f} leaf then refuses to close, and $\Wopen$ stays
  near $1$.  A benchmark-aware bootstrap policy---draw until at least
  one hit is observed in every reachable branch---would unblock this
  regime.
\end{enumerate}

\paragraph{Future work.}
Two extensions deserve immediate attention.  First, replacing the
candidate-clause scorer with Craig
interpolation~\cite{mcmillan2003interpolation,henzinger2004abstraction}
would let the algorithm synthesise predicates the program never
explicitly branches on, broadening the class of programs for which
$\Wopen$ can be driven to zero.  Second, substituting a certified
approximate model counter~\cite{kiesl2024certified,yang2023approxmc6}
for the importance-sampling estimator on non-axis-aligned regions
would yield end-to-end machine-checkable certificates.  Beyond these,
we are interested in (a)~extending the distribution class to (D2)
Bayes-net-structured inputs, (b)~parallel SMT to mitigate the
single-threaded closure-cost bottleneck, and (c)~head-to-head
comparison against exact-inference tools such as PSI~\cite{psi} and
Dice~\cite{holtzen2020dice} on benchmarks ported to their respective
input languages.

\paragraph{Reproducibility.}
\toolname{}, including all benchmarks, scripts, the Docker image, and
the experimental study underlying \Cref{sec:results}, is anonymised
for review at \code{https://github.com/<anon>/dise}.  All tables and
figures are reproduced end-to-end by
\code{experiments/run\_full\_study.py} and
\code{experiments/run\_rare\_event\_scaling.py}, followed by
\code{experiments/plot\_results.py}.
